---
title: "p8105_hw2_sm5134"
author: "Sneha Mehta"
date: "2022-10-04"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1: Cleaning NYC Transit Data
Importing and viewing the dataset:
```{r}
nyc_transit = read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

tail(nyc_transit)
skimr::skim(nyc_transit)
```

This dataset describes characteristics of subways stations in NYC. Key variables include station name, location (longitude and latitude), entrance type, exit locations, ADA accessibility, and routes serviced. This data has been cleaned so far by cleaning variable names, selecting variables of interest, and transforming variables from character to logical for clarity. The resulting dataset has 1868 observations and 20 variables. This dataset is not yet tidy. It would be clearer to transpose the data into a long format.



How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
```{r}
nyc_transit %>% 
  select(station_name, line) %>% 
  distinct
```
There are 465 distinct stations in this dataset.

How many stations are ADA compliant?
```{r}
nyc_transit %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
84 stations are ADA compliant. 

What proportion of station entrances / exits without vending allow entrance? 
```{r}
nyc_transit %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
37.7% of the station entrances/exits without vending allow entrance.

## Problem 2: Mr. Trash Wheel
```{r mrtrash}
mrtrash = read_excel("./Trash Wheel Collection Data.xlsx", range = "Mr. Trash Wheel!A2:N549") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = as.integer(sports_balls)) %>% mutate(wheel = "mister")
```

Repeating with Professor Trash Wheel
```{r proftrash}
proftrash = read_excel("./Trash Wheel Collection Data.xlsx", range = "Professor Trash Wheel!A2:M96") %>% 
  janitor::clean_names() %>% mutate(wheel = "professor")%>% mutate(year = as.character(year))
```

Joining mrtrash and proftrash data
```{r join}
trashjoin = bind_rows(mrtrash, proftrash) %>% 
  janitor::clean_names() 
```

Data Observations
```{r}
skimr::skim(trashjoin)
summary(trashjoin)
sum(trashjoin$weight_tons)
```
There are 641 observations (dumpsters) with 15 variables in the resulting dataset. Key variables in this dataset include weight of trash gathered (in tons) and amounts of the various types of trash gathered (including but not limited to chip bags, plastic bottles, cigarette butts, polystyrene, etc.) Both trash wheels gathered a combined total of 1938 tons of trash. Professor Trash Wheel collected `r sum(trashjoin[which(trashjoin$wheel=="professor"), "weight_tons"])` tons of trash. Mr. Trash Wheel gathered `r sum(trashjoin[which(trashjoin$year=="2020" & trashjoin$wheel == "mister"), "sports_balls"])` sports balls in 2020. 


##Problem 3:FiveThirtyEight
```{r import and clean pols-month}
pols_month = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(president = case_when(prez_gop == 1 ~ "gop", 
                               prez_dem == 1 ~ "dem")) %>% 
  select(-prez_dem, -prez_gop, -day)

```
```{r import and clean snp}
snp = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE)  %>% 
  mutate(month = month.abb[month]) %>% 
  select(-day) %>%
  mutate(year = ifelse(year > 49, year + 1900, year + 2000)) %>% 
  select("year", "month", "close")
```
```{r unemployment}
unemp = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment") %>% 
  mutate(month = str_to_title(month))
```
```{r merging datasets}
snp_pols = left_join(pols_month, snp)
merged_data = left_join(snp_pols, unemp, by = c("year", "month"))
summary(merged_data$year)
```
This dataset describes several measures of politics and the economy in the United States across several election years. The dataset has 822 observations and 11 variables. Key variables include election date (year and month) ranging from 1947-2015, unemployment rate on the year and month of the election, closing value of the S&P stock index on the election date, number of government representatives in office on the date of the election by political party.
