---
title: "p8105_hw2_sm5134"
author: "Sneha Mehta"
date: "2022-10-04"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1: Cleaning NYC Transit Data
Importing and viewing the dataset:
```{r}
nyc_transit = read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
tail(nyc_transit)
skimr::skim(nyc_transit)
janitor::clean_names(nyc_transit) %>% select(line, station_name, station_latitude, station_longitude, starts_with("route"), entrance_type, entry, exit_only, ada) %>% mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```


Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance? 

## Problem 2: Mr. Trash Wheel
```{r mrtrash}
mrtrash = read_excel("./Trash Wheel Collection Data.xlsx", range = "Mr. Trash Wheel!A2:N549") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

Repeating with Professor Trash Wheel
```{r proftrash}
proftrash = read_excel("./Trash Wheel Collection Data.xlsx", range = "Professor Trash Wheel!A2:M96") %>% 
  janitor::clean_names()
```

